### 2023-09-29 manuscript.tex
update texts [ci skip]
This build is based on
https://github.com/MMV-Lab/im2im-paper/commit/805720c4c9ab45e392cf7c613f4c525f9a8c0b2c.

This commit was created by the following CI build and job:
https://github.com/MMV-Lab/im2im-paper/commit/805720c4c9ab45e392cf7c613f4c525f9a8c0b2c/checks
https://github.com/MMV-Lab/im2im-paper/actions/runs/6357857341

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/70ca43307d46352195ff4a7e7a8292144b340afb/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/70ca43307d46352195ff4a7e7a8292144b340afb}{MMV-Lab/im2im-paper@70ca433}
on September 28, 2023.

@@

\begin{itemize}
\item
  \textbf{Justin Sonneck}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-1640-3045}{0000-0002-1640-3045}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/Justin-Sonneck}{Justin-Sonneck}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JustinSonneck}{JustinSonneck}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\item
  \textbf{Jianxu Chen}
  \textsuperscript{\protect\hyperlink{correspondence}{✉}}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-8500-1357}{0000-0002-8500-1357}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/jxchen01}{jxchen01}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JianxuChen}{JianxuChen}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\end{itemize}

@@

Over the past decade, deep learning (DL) research in computer vision has been growing rapidly, with many advances in DL-based image analysis methods for biomedical problems. In this work, we introduce MMV\_Im2Im, a new open-source python package for image-to-image transformation in bioimaging applications. MMV\_Im2Im is designed with a generic image-to-image transformation framework that can be used for a wide range of tasks, including semantic segmentation, instance segmentation, image restoration, and image generation, etc.. Our implementation takes advantage of state-of-the-art machine learning engineering techniques, allowing researchers to focus on their research without worrying about engineering details. We demonstrate the effectiveness of MMV\_Im2Im on more than ten different biomedical problems, showcasing its general potentials and applicabilities.

For computational biomedical researchers, MMV\_Im2Im provides a starting point for developing new biomedical image analysis or machine learning algorithms, where they can either reuse the code in this package or fork and extend this package to facilitate the development of new methods. Experimental biomedical researchers can benefit from this work by gaining a comprehensive view of the image-to-image transformation concept through diversified examples and use cases. We hope this work can give the community inspirations on how DL-based image-to-image transformation can be integrated into the assay development process, enabling new biomedical studies that cannot be done only with traditional experimental assays. To help researchers get started, we have provided source code, documentation, and tutorials for MMV\_Im2Im at https://github.com/MMV-Lab/mmv\_im2im under MIT license.

@@

With the rapid advancements in the fields of machine learning (ML) and computer vision, computers can now transform images into new forms, enabling better visualization {[}\protect\hyperlink{ref-1O0bopKD}{1}{]}, better animation {[}\protect\hyperlink{ref-LxlUp436}{2}{]} and better information extraction {[}\protect\hyperlink{ref-11chATuF4}{3}{]} with unprecedented and continuously growing accuracy and efficiency compared to conventional digital image processing.
These techniques have recently been adapted for bioimaging applications and have revolutionized image-based biomedical research {[}\protect\hyperlink{ref-Yq8wZ6hc}{4},\protect\hyperlink{ref-WwenuBHa}{5},\protect\hyperlink{ref-wcCVn8av}{6},\protect\hyperlink{ref-xPgDok51}{7}{]}. In principal, these techniques and applications can be formulated as a general image-to-image transformation problem, as depicted in the central panel in Figure \ref{fig:overview}. Deep neural networks are trained to perceive the information from the source image(s) and reconstruct the learned knowledge from source images(s) in the form of a new image(s) of the target type.
The source and target images can be real or simulated microscopy images, segmentation masks, or their combinations, as exemplified in Figure \ref{fig:overview}.
Since these underlying methods share the same essential spirit, a natural question arises: is it possible to develop a single generic codebase for deep learning (DL) based image-to-image transformation applicable to various biomedical studies?

@@

Why do we need such a single generic codebase for all deep-learning based microscopy image-to-image transformation? \emph{MMV\_Im2Im} is not simply a collection of many existing methods, but with rather systematic design for generality, flexibility, simplicity and reusability, attempting to address several fundamental bottlenecks for image-to-image transformation in biomedical applications, as highlighted below.

@@

Moreover, as ML scientists, have you ever overwhelmed by different training tricks for different methods or been curious about if certain state-of-the-art training methods can boost the performance of existing models? With the pytorch-lightning backend, \emph{MMV\_Im2Im} allows you to enjoy different state-of-the-art ML training engineering techniques without changing the code, e.g., stochastic weight averaging {[}\protect\hyperlink{ref-qzeQFRn9}{15}{]}, single precision training, automatic batch size determination, different optimizers, different learning rate schedulers, easy deployment on different devices, distributed training on multi-GPU (even multi-node), logging with common loggers such as Tensorboard, etc.. In short, with the pytorch-lightning based universal boilerplate, bioimaging researchers can really focus on research and develop novel methods for their biomedical applications, without worrying about the ML engineering works (which are usually lack in non-computer-science labs).

@@

The toolbox is designed for both people with or without extensive experience with ML and Python programming. It is not rare to find biomedical image analysis software that are very easy to use on a set of problems, but very hard to extend or adjust to other different but essentally related problems, or find some with great flexibility with tunable knobs at all levels, but unfortunately not easy for unexperienced users. To address this issue, we design the toolbox in a systematically modularized way with various levels of configurability. One can use the toolbox with a single command as simple as \texttt{run\_im2im\ -\/-config\ train\_semanticseg\_3d\ -\/-data.data\_path\ /path/to/data} or make customization on details directly from a human-readable configuration file, such as choosing batch normalization or instance normalization in certain layers of the model, or adding extra data augmentation steps, etc.. We provide an extensive list of more than 20 example configurations for various appplications and comprehensive documentation to address common questions for users as reference. For users preferring graphical interface, another napari plugin for MMV toolbox has been planned as the extension of \emph{MMV\_Im2Im} (see Discussion for details).

@@

The original idea of a general toolbox actually stememd from the OpenMMLab project (\url{https://openmmlab.com/}), which provides generic codebases for a wide range of computer vision research topics. For instance, \emph{MMSegmentation} (\url{https://github.com/open-mmlab/mmsegmentation}) is an open source toolbox for semantic segmentation, supporting unified benchmarking and state-of-the-art models ready to use out-of-box.
It has become one of most widely used codebase for research in semantic segmentation (2K forks and 5.4K stars on Github as of March 12, 2023).
This inspires us to develop \emph{MMV\_Im2Im} to facillitate research in image-to-image transformation with special focus on biomedical applications.

First of all, different from general computer vision datasets, such as ImageNet {[}\protect\hyperlink{ref-lt4BNUoG}{16}{]}, where the images are usually small 2D RGB images (e.g., 3 x 256 x 256 pixels), biomedical applications usually involves large-scale high dimensional data (e.g., 500 images of 4 x 128 x 2048 x 2048 voxels). To deal with this issue, we employ the PersistentDataset in MONAI {[}\protect\hyperlink{ref-UU62HYC6}{17}{]} with partial loading and sampling support, as well as delayed image reading powered by aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{18}{]} as default (configurable if another dataloader is preferred).
As a result, in our stress test, training a 3D nuclei instance segmentation model with more than 125,000 3D images can be conducted efficiently in a day, even with limited resource.

@@

Third, the toolbox pre-packs common functionalities specific to microscopy images. For example, we incorporate the special image normalization method introduced in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, where only the middle chunk along Z dimension of 3D microscopy images will be used for calculating the mean and standard deviation of image intensity for standard normalization. Also, 3D light microscopy images are usually anisotropic, i.e., much lower resolution along Z than XY dimension. So, we adopt the anisotropic variation of UNet as proposed in {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]}.

Finally, to deploy the model in production, a model trained on small 3D patches sometimes need to be applied not only on much large images. Combining the efficient data handling of aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{18}{]} and the sliding window inference with gaussian weighted blending, the toolbox can yield efficient inference without visible stitching artifacts in production.

All in all, the \emph{MMV\_Im2Im} toolbox stands on the shoulders of many giants in the open-source software and ML engineering communities (pytorch-lightning, MONAI, aicsimageio, etc.) and is systematically designed for image-to-image transformation R\&D for biomedical applications. The source code of \emph{MMV\_Im2Im} is available at \url{https://github.com/MMV-Lab/mmv_im2im}. This manuscript is generated with open-source package Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{20}{]}. The manuscript source code is available at \url{https://github.com/MMV-Lab/im2im-paper}.

@@

In this section, we showcase the versatility of the \emph{MMV\_Im2Im} toolbox by presenting over ten different biomedical applications across various R\&D use cases and scales. All experiments and results in this section were conducted on publicly available datasets released with other publications and our scripts (for pulling the public dataset online and data wrangling) and configuration files (for setting up training and inference details), both included in the MMV\_Im2Im package. Our aim is to make it easy to reproduce all of the results in this paper, and more importantly use these data and scripts to get familar with the package and adapt to new problems of users' interest. It is important to note that the aim of these experiments was not to achieve the best performance on each individual task, as this may require further hyper-parameter tuning (see Discussion section for more details). Rather, the experiments were intended to demonstrate the package's different features and general applicability, providing a holistic view of image-to-image transformation concepts to biomedical researchers. We hope that these concepts will help researchers integrate AI into traditional assay development strategies and inspire computational and experimental co-design methods, enabling new biomedical studies that were previously unfeasible.

\hypertarget{labelfree-prediction-of-nuclear-structure-from-2d3d-brightfield-imagestodo-rename-unet-to-unet-in-fig-a}{%
\subsubsection{Labelfree prediction of nuclear structure from 2D/3D brightfield images(TODO: rename UNET to UNet in Fig A)}\label{labelfree-prediction-of-nuclear-structure-from-2d3d-brightfield-imagestodo-rename-unet-to-unet-in-fig-a}}

The labelfree method refers a DL method that can predict fluorescent images directly from transmitted light brightfield images {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. Comparing to brightfield images, fluorescent images can resolve subcellular structures in living cells at high resolution but with the cost of expensive and slow procedures and high phototoxicity. The labelfree method provides a new perspective in assay development to conduct integrated computational analysis of multiple organelles only with a single brightfield image acquisition. In our first demonstration, we applied \emph{MMV\_Im2Im} to build 2D/3D models that can predict fluorescent images of nuclear structures from brightfield images. For 3D models, we also compared (1) different image normalization methods, (2) different network backbones, and (3) different types of models.

@@

\emph{2D Labelfree:} We started with a simple problem using 2D images from the HeLa ``Kyoto'' cells dataset {[}\protect\hyperlink{ref-xv2VIyRP}{23}{]}. For all images, we took the brightfield channel and the mCherry-H2B channel out of the multi-channel timelapse movies. 2D images were acquired at 20x with 0.8 N.A. and then downscaled by 4 (pixel size: 0.299 nm x 0.299 nm). Example predictions can be found in Figure \ref{fig:labelfree}-A. We compared a basic UNet model {[}\protect\hyperlink{ref-TutLhFSz}{9}{]} and a 2D version of the fnet model in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. The fnet model achieved slightly more accurate predictions than the basic UNet, this can be seen in Figure \ref{fig:labelfree_comparison}-A.

@@

\begin{figure}
\hypertarget{fig:labelfree_comparison}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree_comparison_justin.png}
\caption{A. Example of 2D labelfree results. B. Overview of various 3D labelfree results obtained by different training strategies. p/c/s refers percentile normalization, center normalization, and standard normalization, respectively (see main text for details).}\label{fig:labelfree_comparison}
}
\end{figure}

We conducted three groups of comparisons (see results in Figure \ref{fig:labelfree_comparison}-B). First, we compared three different image normalization methods for 3D images: percentile normalization, standard normalization, center normalization {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. Percentile normalization refers to cutting the intensity out of the range of {[}0.5, 99.5{]} percentile of the image intensity and then rescale the values to the range of {[}-1, 1{]}, while the standard normalization is simply subtracting mean intensity and then divided by the standard deviation of all pixel intensities. Center normalization is similar to standard normalization, but the statistics are calculated only around center along the Z-axis {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. One could easily test different percentile or rescaling to {[}0, 1{]} instead of {[}-1, 1{]}. Qualitatively, we found center normalization slightly more accurate and more robust than the other two (ref. the first row in Figure \ref{fig:labelfree}-B).

Second, we compared different network backbone architectures, including the original fnet model {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, an enhanced UNet {[}\protect\hyperlink{ref-M7480NLD}{25}{]}, the attention UNet {[}\protect\hyperlink{ref-OCow1hly}{26}{]}, two transformer-based models, SwinUNETR {[}\protect\hyperlink{ref-ZWL3IrVc}{27}{]} and UNETR{[}\protect\hyperlink{ref-XCKUntOB}{28}{]} (all with center normalization). Inspecting the predictions on a holdout validation set suggested that fnet achieved the best performance \#, and the transformer-based models did not work well in labelfree problems (ref. the second row and the ``c + fnet'' from the first row in Figure \ref{fig:labelfree}-B).

Finally, we showed the comparison between three different types of models, an FCN-type model (i.e., fnet), a pix2pix-type model, and a cycleGAN-type model. For fair comparison, we used fnet as the same backbone for all three types of models. In theory, the pix2pix-type model can be trained in two different ways: from scratch or initializing the generator with a pre-trained fnet (trained as FCN). Examples of the comparison results were shown in the last two rows in Figure \ref{fig:labelfree}-B. Visually, it is evident that the additional adversarial components (i.e., the discriminator) could generate images with more realistic appearance than a typical FCN-type model alone, but again, we leave the appropriate quantitative evaluations to users' specfic biomedical studies.

From the experiments above, we found that center normalization + pix2pix with fnet as the generator achieved the best overall performance qualitatively. So, we employed the same strategy on all other nuclear structures. At the end, we had four different labelfree models, each predicting one different nuclear structure from 3D brightfield images. As an example of evaluation, we calculated the pearson correlation, the structural similarity and the peak signal to noise ratio on hold-out validation sets. The results were summarized in Table \ref{tbl:labelfree_table}. Again, these numbers were merely examples of evaluation, systematic evaluation based on each specific biological problem would be necessary before deployment. Figure \ref{fig:labelfree_all_structures}-A shows the comparison of each predicted structure and its ground truth, while Figure \ref{fig:labelfree_all_structures}-B shows one example of all four different structures predicted from a single unseen brightfield image. This would permit an integrated analysis of four different nuclear components that could hardly be acquired simultaneously in real experiments and real images.

@@

\begin{figure}
\hypertarget{fig:labelfree_all_structures}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree_all_justin.png}
\caption{A. Comparison of predictions of different 3d labelfree models and each ground truth. B. Predictions of the different labelfree models using the same brightfield image as input, which provides a deep insight into the nuclear structure.}\label{fig:labelfree_all_structures}
}
\end{figure}

@@

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. DL-based methods have achieved huge success in semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{29},\protect\hyperlink{ref-XAffSYIR}{30}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{31}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{26}{]}. Evaluated on the two different hold-out test sets, the model achieved F1-score, 0.904 ± 0.060 and 0.861 ± 0.117 on test set A and test set B, respectively. The performance was competitive compared to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{30}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.

>>>>

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/805720c4c9ab45e392cf7c613f4c525f9a8c0b2c/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/805720c4c9ab45e392cf7c613f4c525f9a8c0b2c}{MMV-Lab/im2im-paper@805720c}
on September 29, 2023.

@@

\begin{itemize}
\item
  \textbf{Justin Sonneck}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-1640-3045}{0000-0002-1640-3045}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/Justin-Sonneck}{Justin-Sonneck}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JustinSonneck}{JustinSonneck}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\item
  \textbf{Yu Zhou}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0009-0002-3914-1102}{0009-0002-3914-1102}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/audreyeternal}{audreyeternal}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/eternalaudreys}{eternalaudreys}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\item
  \textbf{Jianxu Chen}
  \textsuperscript{\protect\hyperlink{correspondence}{✉}}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-8500-1357}{0000-0002-8500-1357}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/jxchen01}{jxchen01}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JianxuChen}{JianxuChen}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\end{itemize}

@@

Over the past decade, deep learning (DL) research in computer vision has been growing rapidly, with many advances in DL-based image analysis methods for biomedical problems. In this work, we introduce \emph{MMV\_Im2Im}, a new open-source python package for image-to-image transformation in bioimaging applications. \emph{MMV\_Im2Im} is designed with a generic image-to-image transformation framework that can be used for a wide range of tasks, including semantic segmentation, instance segmentation, image restoration, and image generation, etc.. Our implementation takes advantage of state-of-the-art machine learning engineering techniques, allowing researchers to focus on their research without worrying about engineering details. We demonstrate the effectiveness of \emph{MMV\_Im2Im} on more than ten different biomedical problems, showcasing its general potentials and applicabilities.

For computational biomedical researchers, \emph{MMV\_Im2Im} provides a starting point for developing new biomedical image analysis or machine learning algorithms, where they can either reuse the code in this package or fork and extend this package to facilitate the development of new methods. Experimental biomedical researchers can benefit from this work by gaining a comprehensive view of the image-to-image transformation concept through diversified examples and use cases. We hope this work can give the community inspirations on how DL-based image-to-image transformation can be integrated into the assay development process, enabling new biomedical studies that cannot be done only with traditional experimental assays. To help researchers get started, we have provided source code, documentation, and tutorials for \emph{MMV\_Im2Im} at https://github.com/MMV-Lab/mmv\_im2im under MIT license.

@@

With the rapid advancements in the fields of machine learning (ML) and computer vision, computers can now transform images into new forms, enabling better visualization {[}\protect\hyperlink{ref-1O0bopKD}{1}{]}, better animation {[}\protect\hyperlink{ref-LxlUp436}{2}{]} and better information extraction {[}\protect\hyperlink{ref-11chATuF4}{3}{]} with unprecedented and continuously growing accuracy and efficiency compared to conventional digital image processing. These techniques have recently been adapted for bioimaging applications and have revolutionized image-based biomedical research {[}\protect\hyperlink{ref-Yq8wZ6hc}{4},\protect\hyperlink{ref-WwenuBHa}{5},\protect\hyperlink{ref-wcCVn8av}{6},\protect\hyperlink{ref-xPgDok51}{7}{]}. In principle, these techniques and applications can be formulated as a general image-to-image transformation problem, as depicted in the central panel in Figure 1. Deep neural networks are trained to perceive the information from the source image(s) and reconstruct the learned knowledge from source images(s) in the form of a new image(s) of the target type. The source and target images can be real or simulated microscopy images, segmentation masks, or their combinations, as exemplified in Figure 1. Since these underlying methods share the same essential spirit, a natural question arises: is it possible to develop a single generic codebase for deep learning (DL) based image-to-image transformation applicable to various biomedical studies?

@@

Why do we need such a single generic codebase for all deep-learning based microscopy image-to-image transformation? \emph{MMV\_Im2Im} is not simply a collection of many existing methods, but rather has a systematic design for generality, flexibility, simplicity and reusability, attempting to address several fundamental bottlenecks for image-to-image transformation in biomedical applications, as highlighted below.

@@

Moreover, as ML scientists, have you ever been overwhelmed by different training tricks for different methods or been curious about if certain state-of-the-art training methods can boost the performance of existing models? With the pytorch-lightning backend, \emph{MMV\_Im2Im} allows you to enjoy different state-of-the-art ML training engineering techniques without changing the code, e.g., stochastic weight averaging {[}\protect\hyperlink{ref-qzeQFRn9}{15}{]}, single precision training, automatic batch size determination, different optimizers, different learning rate schedulers, easy deployment on different devices, distributed training on multi-GPU (even multi-node), logging with common loggers such as Tensorboard, etc.. In short, with the pytorch-lightning based universal boilerplate, bioimaging researchers can really focus on research and develop novel methods for their biomedical applications, without worrying about the ML engineering works (which are usually lacking in non-computer-science labs).

@@

The toolbox is designed for both people with or without extensive experience with ML and Python programming. It is not rare to find biomedical image analysis software that is very easy to use on a set of problems, but very hard to extend or adjust to other different but essentially related problems, or find some with great flexibility with tunable knobs at all levels, but unfortunately not easy for inexperienced users. To address this issue, we design the toolbox in a systematically modularized way with various levels of configurability. One can use the toolbox with a single command as simple as \texttt{run\_im2im\ -\/-config\ train\_semanticseg\_3d\ -\/-data.data\_path\ /path/to/data} or make customization on details directly from a human-readable configuration file, such as choosing batch normalization or instance normalization in certain layers of the model, or adding extra data augmentation steps, etc.. We provide an extensive list of more than 20 example configurations for various applications and comprehensive documentation to address common questions for users as reference. For users preferring graphical interface, another napari plugin for MMV toolbox has been planned as the extension of \emph{MMV\_Im2Im} (see Discussion for details).

@@

The original idea of a general tool box actually stemmed from the OpenMMLab project (\url{https://openmmlab.com/}), which provides generic codebases for a wide range of computer vision research topics. For instance, \emph{MMSegmentation} (\url{https://github.com/open-mmlab/mmsegmentation}) is an open source toolbox for semantic segmentation, supporting unified benchmarking and state-of-the-art models ready to use out-of-box. It has become one of most widely used codebase for research in semantic segmentation (2.3K forks and 6.5K stars on Github as of September 29, 2023). This inspires us to develop \emph{MMV\_Im2Im} to facilitate research in image-to-image transformation with special focus on biomedical applications.

First of all, different from general computer vision datasets, such as ImageNet {[}\protect\hyperlink{ref-lt4BNUoG}{16}{]}, where the images are usually small 2D RGB images (e.g., 3 x 256 x 256 pixels), biomedical applications usually involves large-scale high dimensional data (e.g., 500 images of 4 x 128 x 2048 x 2048 voxels). To deal with this issue, we employ the PersistentDataset in MONAI {[}\protect\hyperlink{ref-UU62HYC6}{17}{]} with partial loading and sampling support, as well as delayed image reading powered by aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{18}{]} as default (configurable if another dataloader is preferred). As a result, in our stress test, training a 3D nuclei instance segmentation model with more than 125,000 3D images can be conducted efficiently in a day, even with limited resources.

@@

Third, the toolbox pre-packs common functionalities specific to microscopy images. For example, we incorporate the special image normalization method introduced in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, where only the middle chunk along the Z dimension of 3D microscopy images will be used for calculating the mean and standard deviation of image intensity for standard normalization. Also, 3D light microscopy images are usually anisotropic, i.e., much lower resolution along Z than XY dimension. So, we adopt the anisotropic variation of UNet as proposed in {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]}.

Finally, to deploy the model in production, a model trained on small 3D patches sometimes needs to be applied not only on much larger images. Combining the efficient data handling of aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{18}{]} and the sliding window inference with gaussian weighted blending, the toolbox can yield efficient inference without visible stitching artifacts in production.

All in all, the \emph{MMV\_Im2Im} toolbox stands on the shoulders of many giants in the open-source software and ML engineering communities (pytorch-lightning, MONAI, aicsimageio, etc.) and is systematically designed for image-to-image transformation R\&D for biomedical applications. The source code of \emph{MMV\_Im2Im} is available at \url{https://github.com/MMV-Lab/mmv_im2im}. This manuscript is generated with the open-source package Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{20}{]}. The manuscript source code is available at \url{https://github.com/MMV-Lab/im2im-paper}.

@@

In this section, we showcase the versatility of the \emph{MMV\_Im2Im} toolbox by presenting over ten different biomedical applications across various R\&D use cases and scales. All experiments and results in this section were conducted on publicly available datasets released with other publications and our scripts (for pulling the public dataset online and data wrangling) and configuration files (for setting up training and inference details), both included in the \emph{MMV\_Im2Im} package. Our aim is to make it easy to reproduce all of the results in this paper, and more importantly use these data and scripts to get familiar with the package and adapt to new problems of users' interest. It is important to note that the aim of these experiments was not to achieve the best performance on each individual task, as this may require further hyper-parameter tuning (see Discussion section for more details). Rather, the experiments were intended to demonstrate the package's different features and general applicability, providing a holistic view of image-to-image transformation concepts to biomedical researchers. We hope that these concepts will help researchers integrate AI into traditional assay development strategies and inspire computational and experimental co-design methods, enabling new biomedical studies that were previously unfeasible.

\hypertarget{labelfree-prediction-of-nuclear-structure-from-2d3d-brightfield-images}{%
\subsubsection{Labelfree prediction of nuclear structure from 2D/3D brightfield images}\label{labelfree-prediction-of-nuclear-structure-from-2d3d-brightfield-images}}

The labelfree method refers to a DL method that can predict fluorescent images directly from transmitted light brightfield images {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. Compared to brightfield images, fluorescent images can resolve subcellular structures in living cells at high resolution but with the cost of expensive and slow procedures and high phototoxicity. The labelfree method provides a new perspective in assay development to conduct integrated computational analysis of multiple organelles only with a single brightfield image acquisition. In our first demonstration, we applied \emph{MMV\_Im2Im} to build 2D/3D models that can predict fluorescent images of nuclear structures from brightfield images. For 3D models, we also compared (1) different image normalization methods, (2) different network backbones, and (3) different types of models.

@@

\emph{2D Labelfree:} We started with a simple problem using 2D images from the HeLa ``Kyoto'' cells dataset {[}\protect\hyperlink{ref-xv2VIyRP}{23}{]}. For all images, we took the brightfield channel and the mCherry-H2B channel out of the multi-channel timelapse movies. 2D images were acquired at 20x with 0.8 N.A. and then downscaled by 4 (pixel size: 0.299 nm x 0.299 nm). Example predictions can be found in Figure \ref{fig:labelfree_comparison}-A. We compared a basic UNet model {[}\protect\hyperlink{ref-TutLhFSz}{9}{]} and a 2D version of the fnet model in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. The fnet model achieved slightly more accurate predictions than the basic UNet, this can be seen in Figure \ref{fig:labelfree_comparison}-A.

@@

\begin{figure}
\hypertarget{fig:labelfree_comparison}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree_comparison_justin.png}
\caption{A. Example of 2D labelfree results. B. Overview of various 3D labelfree results obtained by different training strategies. p/c/s refers to percentile normalization, center normalization, and standard normalization, respectively (see main text for details). (The contrast of grayscale images was adjusted using ImageJ's autoscale.)}\label{fig:labelfree_comparison}
}
\end{figure}

We conducted three groups of comparisons (see results in Figure \ref{fig:labelfree_comparison}-B). First, we compared three different image normalization methods for 3D images: percentile normalization, standard normalization, center normalization {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. Percentile normalization refers to cutting the intensity out of the range of {[}0.5, 99.5{]} percentile of the image intensity and then rescale the values to the range of {[}-1, 1{]}, while the standard normalization is simply subtracting mean intensity and then divided by the standard deviation of all pixel intensities. Center normalization is similar to standard normalization, but the statistics are calculated only around the center along the Z-axis {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}. One could easily test different percentiles or rescaling to {[}0, 1{]} instead of {[}-1, 1{]}. Qualitatively, we found center normalization slightly more accurate and more robust than the other two (see first row of Figure \ref{fig:labelfree}-B).

Second, we compared different network backbone architectures, including the original fnet model {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, an enhanced UNet {[}\protect\hyperlink{ref-M7480NLD}{25}{]}, the attention UNet {[}\protect\hyperlink{ref-OCow1hly}{26}{]}, two transformer-based models, SwinUNETR {[}\protect\hyperlink{ref-ZWL3IrVc}{27}{]} and UNETR{[}\protect\hyperlink{ref-XCKUntOB}{28}{]} (all with center normalization). Inspecting the predictions on a holdout validation set suggested that fnet achieved the best performance.

Finally, we showed the comparison between three different types of models, an FCN-type model (i.e., fnet), a pix2pix-type model, and a cycleGAN-type model. For fair comparison, we used fnet as the same backbone for all three types of models. In theory, the pix2pix-type model can be trained in two different ways: from scratch or initializing the generator with a pre-trained fnet (trained as FCN). Examples of the comparison results were shown in the last two rows in Figure \ref{fig:labelfree}-B. Visually, it is evident that the additional adversarial components (i.e., the discriminator) could generate images with more realistic appearance than a typical FCN-type model alone, but again, we leave the appropriate quantitative evaluations to users' specific biomedical studies.

From the experiments above, we found that center normalization + pix2pix with fnet as the generator achieved the best overall performance qualitatively. So, we employed the same strategy on all other nuclear structures. At the end, we had four different labelfree models, each predicting one different nuclear structure from 3D brightfield images. As an example of evaluation, we calculated the pearson correlation, the structural similarity and the peak signal to noise ratio on holdout validation sets. The results were summarized in Table \ref{tbl:labelfree_table}. Again, these numbers were merely examples of evaluation, systematic evaluation based on each specific biological problem would be necessary before deployment. Figure \ref{fig:labelfree_all_structures}-A shows the comparison of each predicted structure and its ground truth, while Figure \ref{fig:labelfree_all_structures}-B shows one example of all four different structures predicted from a single unseen brightfield image. This would permit an integrated analysis of four different nuclear components that could hardly be acquired simultaneously in real experiments and real images.

@@

\begin{figure}
\hypertarget{fig:labelfree_all_structures}{%
\centering
\includegraphics[width=0.76\textwidth,height=0.9\textheight]{images/labelfree_all_justin.png}
\caption{A. Comparison of predictions of different 3D labelfree models and each ground truth. B. Predictions of the different labelfree models using the same brightfield image as input, which provides a deep insight into the nuclear structure. This would not be possible with brightfield imaging alone and is enabled by the application of the labelfree approach. (The contrast of grayscale images was adjusted using ImageJ's autoscale.)}\label{fig:labelfree_all_structures}
}
\end{figure}

@@

Segmentation is a common image processing task, and can be considered as a special type of image-to-image transformation, where the generated images are segmentation masks. DL-based methods have achieved huge success in semantic segmentation in biomedical images. In this example, we demonstrated \emph{MMV\_Im2Im} on a pathology application to segment glands from hematoxylin and eosin (H\&E) stained tissue images from the 2015 Gland Segmentation challenge {[}\protect\hyperlink{ref-45Sirz1X}{29},\protect\hyperlink{ref-XAffSYIR}{30}{]}. Stain normalization is an important pre-processing step in order to develop models robust to stain variation and tissue variations. \emph{MMV\_Im2Im} included a classic stain normalization method {[}\protect\hyperlink{ref-tQhnZyjK}{31}{]} as a pre-processing step. The effect of stain normalization can be observed in Figure \ref{fig:2d_gland}-A and B. We trained a simple attention UNet model {[}\protect\hyperlink{ref-OCow1hly}{26}{]}. Evaluated on the two different holdout test sets, the model achieved F1-score, 0.904 ± 0.060 and 0.861 ± 0.117 on test set A and test set B, respectively. The performance was competitive compared to the methods reported in the challenge report {[}\protect\hyperlink{ref-XAffSYIR}{30}{]}, especially with much more consistent performance across the two different test sets. Example results can be found in Figure \ref{fig:2d_gland}-C.
